On the probability of finding nonphysical solutions through shadowing\tnoteref{1}

### 2021-04-24 letter/responseToReviewers.tex

1. \reviewerOne{There are some published articles that support some conclusions of this manuscript...All of these should give the authors more information about the topic related to this manuscript}

We thank the reviewer for pointing out multiple references that support the conclusions of the present paper. In our revised manuscript, we have added them to our discussion section 5 (the edits are highlighted in \reviewerOne{red}), where we had originally referred to Clean Numerical Simulation (CNS).

2. \reviewerOne{There is another thing that might be mentioned.  Periodic solutions of some chaotic systems, such as three-body system, have physical meanings.  Although three-body systems are chaotic in essence, periodic orbits of three-body system indeed exist.}

Thank you for pointing this important point out. In our paper, we use ``nonphysicality'' to mean ``atypicality'' in the following sense, which, as you point out, may not be the same as physically meaningless. Whenever an initial condition is sampled from a set of full Lebesgue measure in an open set, known as the attractor basin, that contains the chaotic attractor, the ergodic average along the orbit starting from that initial condition is equal to the ensemble average. This latter ensemble average or phase-space average is with respect to an SRB-type probability measure, which is called a {\em physical} measure (Ref. 10) since it is ``observed'' through ergodic averages along almost every orbit starting in the attractor basin. Thus, we refer to as ``nonphysical'' atypical orbits that do not produce the correct statistics associated with the SRB measure.

@@

For an $\epsilon >0$, an $\epsilon$-orbit or a {\em pseudo-orbit} of the original system (Eq. \ref{eqn:true}) is any solution $\hat{x}(t)$ of Eq. \ref{eqn:numerical} wherein $\|\delta f(\hat{x}(t), s)\| \leq \epsilon$ for all $t$. Any reasonably accurate numerical solution can be considered to be an $\epsilon$-orbit of the ground truth dynamics in Eq. \ref{eqn:true}.

This also holds even when a numerical solution may not be modelled by Eq. \ref{eqn:numerical} because $\epsilon$-orbits could be defined more generally without recourse to the particular form of Eq. \ref{eqn:numerical}. Any curve $\hat{x}(t)$ such that $\|\hat{x}'(t) - x'(\hat{x}(t))\| < \epsilon$ for all $t$ is an $\epsilon$-orbit of Eq. \ref{eqn:true}.

The shadowing lemma says that for every $\delta > 0,$ there exists $\epsilon > 0$ such that for every $\epsilon$-orbit $\hat{x}(t)$, there exists a true orbit $x(t)$ satisfying Eq. \ref{eqn:true} that $\delta$-shadows it. This means that there exists some scalar function $\gamma(t)$ with $|\gamma'(t) - 1| < \delta$ such that $\| \hat{x}(\gamma(t)) - x(t) \| < \delta$ for all $t.$

Now, since $x(t)$ and $\hat{x}(t)$ are $\delta$-close to each other, their long-term (empirical) probability distributions are also similar. That is, for any regular scalar function $J,$ the time-averages $\langle J\rangle_x^T := (1/T)\int_0^T J(x(t))\: dt$ and 
$\langle J\rangle_{\hat{x}}^T := (1/T) \int_0^T J(\hat{x}(t)) \: dt$ are close to each other for large $T.$

However, we show through examples in the paper that $x(t)$, a shadowing orbit, which is a true solution of Eq. \ref{eqn:true}, need not be a typical or a {\em physical} orbit. That is, the infinitely-long time average along $x(t)$, $\lim_{T\to\infty} \langle J\rangle_x^T$, need not be equal to the average of $J$ over an ensemble of solutions to Eq. \ref{eqn:true}. In fact, we show in section 5 with the plucked tent map that the discrepancy can be large, when compared to the true value of the ensemble average. Note that the ensemble average over all possible solutions of Eq. \ref{eqn:true} is observed along any randomly chosen orbit, i.e., if we chose an initial condition $x(0)$ by sampling a uniform distribution in the attractor basin of the dynamics in Eq. \ref{eqn:true}, then $\lim_{T\to\infty}\langle J\rangle_x^T$ is equal to the ensemble average of $J.$ However, when $x(0)$ is a special initial condition that is chosen so that $x(t)$ $\delta$-shadows a pseudo-orbit $\hat{x}(t),$ it is not guaranteed to reproduce ensemble statistics.

Since the statistics or long-term averages of pseudo-orbits and shadowing orbits are close, we expect that the pseudoorbit (the numerical solution $\hat{x}(t)$) is also not guaranteed to be statistically consistent.

### 2021-04-24 main_els.tex
Current changes

\reviewerOne{
At the same time, families of periodic solutions have been well-studied for some problems}\cite{henon3body}.\reviewerOne{The set of all unstable periodic orbits have been recognized as the ``skeleton'' of the chaotic attractor \cite{cvitanovic1} forming a dense subset of the attractor and providing an alternative characterization of the SRB distribution \cite{llave}. This implies that ergodic/ensemble averages with respect to the SRB distribution are also weighted averages along the set of periodic orbits.} \reviewerOne{The significance of the set of unstable periodic orbits extends even further as it can be used to compute dynamical invariants such as entropy and Lyapunov dimension \cite{grebogi1}\cite{cvitanovic1}. A quantity of great mathematical interest is the dynamical determinant (see \cite{baladi-book} for a recent review), which is computed as a weighted sum over all periodic orbits. The zeros of the dynamical determinant have a one-to-one correspondence with the slowest decay rates of correlations between observables \cite{baladi-book}. This connection has been exploited both in mathematical as well as physical studies of the distribution of energy levels and resonances in quantum scattering \cite{semyonZworski}\cite{berry}\cite{cvitanovic2005chaos}. Furthermore, theoretical and numerical studies have attributed cataclysmic changes in the shape and position of the attractor in response to small parameter changes, known as attractor crises \cite{grebogi1}, to the emergence of tangencies between the stable and unstable manifolds along periodic orbits.}

>>>>

\reviewerOne{
At the same time, families of periodic solutions have been well-studied for some problems}\cite{henon3body}.\reviewerOne{The set of all unstable periodic orbits have been recognized as the ``skeleton'' of the chaotic attractor \cite{cvitanovic1} forming a dense subset of the attractor and providing an alternative characterization of the SRB distribution \cite{llave}. This implies that ergodic/ensemble averages with respect to the SRB distribution are also weighted averages along the set of periodic orbits.} \reviewerOne{The significance of the set of unstable periodic orbits extends even further as it can be used to compute dynamical invariants such as entropy and Lyapunov dimension \cite{grebogi1}\cite{cvitanovic1}. A quantity of great mathematical interest is the dynamical determinant (see \cite{baladi-book} for a recent review), which is computed as a weighted sum over all periodic orbits. The zeros of the dynamical determinant have a one-to-one correspondence with the slowest decay rates of correlations between observables \cite{baladi-book}. This connection has been exploited both in mathematical as well as physical studies of the distribution of energy levels and resonances in quantum scattering \cite{semyonZworski}\cite{berry}\cite{cvitanovic2005chaos}. Furthermore, theoretical and numerical studies of unstable periodic orbits have been crucial to better understanding transitions to chaos \cite{kreilos} and cataclysmic changes in the shape and position of the attractor in response to small parameter changes, known as attractor crises \cite{grebogi1}\cite{lucariniCrisis}, which are sometimes due to the emergence of tangencies between the stable and unstable manifolds along periodic orbits.}
